Future challenges for Research Software Engineering

### 2025-05-12 future_challenges_for_RSEng.tex
Add side boxes for complexity section

\subsection{Impact of Complexity and Usability in Research Software}
\begin{whatis}{What do we want to have achieved}{abstraction}
Gerasimos? Jan?
\end{whatis}

@@

\begin{story}{The benefits of abstraction}{abstraction}
A story about Kim and Kay effortlessly progressing where years ago they would be caught up in details.
\end{story}

@@

\begin{story}{The greater world of Kim and Kay}{ethics}
10 years have passed since Kim and Kay have started on their journey of being RSEs. Back then,
at the universities of Eden and Utopia, life was more orderly and harmonious. Kay is working in a project that develops
software for cheap water monitoring devices in the global south. Kim is currently designing the software service stack
at her university, and increasingly struggles to find compelling Open Source alternatives that lets them keep their data
to software that is backed by large corporate entities. So far she still has the support of her university board, since it values
strategic autonomy.
Her son Kyle on the other hand started out with a degree in Physics and then took one of the newly created RSE Master's courses,
but quickly decided that work in academia with its time limited contracts is not attractive.
Governments around the world are struggling to reposition themselves in this emerging deglobalized world with its new blocks
and therefore have deprioritized publicly funded research. Kyle decided to go to a defense company since they offer a
more secure employment in this polarized world.
\end{story}

@@

\begin{story}{Changes to the Universities}{changes}
Kim found work years ago at one of the first RSE units in Arcadia. After a couple of years he moved to the university of
Avalon which was interested in setting up their own department for RSEs in order to not fall behind the university of Arcadia.
Kim applies his experiences and improves upon the model set forth by the University of Arcadia. Since the university
of Avalon has to follow different rules not everything can be straightforwardly adapted. Life would be so much easier,
if at least university regulations would be homogenized, no?
\end{story}

>>>>

\subsection{Impact of Complexity and Usability in Research Software}
\begin{whatis}{What do we want to have achieved}{abstraction}
With a palette of readily reusable software components,
and with their skills in managing hierarchical systems of complex software components with clear interfaces,
RSEs can now build systems of scale and capability not previously possible.
\end{whatis}

@@

\begin{story}{The benefits of abstraction}{abstraction}
Years ago, Kim had to distribute the software library libkim as a source code package with a custom build system and detailed instructions on how to build it on some common systems,
investing significant time in developing and maintaining this auxiliary toolkit and documentation.
Around the same time, Kay was developing the research analysis tool kay-bio,
which however only provided an application programming interface that
required writing complex code even for simple and common analysis tasks.
Nowadays, Kim could have started directly from a template based on established toolkits and workflows,
while kay-bio accepts configuration in a simpler, standardized format and generative AI assistants
can guide new users in preparing and debugging their analyses.
\end{story}

@@

\begin{story}{The greater world of Kim and Kay}{ethics}
10 years have passed since Kim and Kay have started on their journey of being RSEs. Back then,
at the universities of Eden and Utopia, life was more orderly and harmonious. Kay is working in a project that develops
software for cheap water monitoring devices in the global south. Kim is currently designing the software service stack
at the university, and increasingly struggles to find compelling Open Source alternatives that lets them keep their data
to software that is backed by large corporate entities. So far she still has the support of the university board, since it values
strategic autonomy.
Her son Kyle on the other hand started out with a degree in Physics and then took one of the newly created RSE Master's courses,
but quickly decided that work in academia with its time limited contracts is not attractive.
Governments around the world are struggling to reposition themselves in this emerging deglobalized world with its new blocks
and therefore have deprioritized publicly funded research. Kyle decided to go to a defense company since they offer a
more secure employment in this polarized world.
\end{story}

@@

\begin{story}{Changes to the Universities}{changes}
Kim found work years ago at one of the first RSE units in Arcadia. After a couple of years he moved to the university of
Avalon which was interested in setting up their own department for RSEs in order to not fall behind the university of Arcadia.
Kim improves upon the model set forth by the University of Arcadia based on valuable experience. Since the university
of Avalon has to follow different rules not everything can be straightforwardly adapted. Life would be so much easier,
if at least university regulations would be homogenized, no?
\end{story}
